Mixture-of-tastes Models for Representing Users with Diverse Interests

### 2017-12-13 paper/paper.tex
Inline results.

\begin{table}
  \caption{Dataset statistics. 95th/50th denotes the ratio of popularity of the 95th and 50th percentile of item popularity.}
  \label{tab:datasets}
  \input{includes/datasets.tex}
\end{table}

@@

\section{Results}
Our main results are summarized in Table \ref{tb:results}. In both sequence-based and factorization tasks, variants of our model achieve consistently higher performance than baseline models, with improvements ranging from 10\% to 69\%. The results are robust to optimizing the hyperparameters of our baselines, giving us ample confidence that the results can be replicated in production recommender systems.
\begin{table}
\caption{Mean reciprocal rank across all users/sequences in the test set. Note that due to differences in experimental protocol, results between sequence-based and factorization models are not directly comparable.}
\label{tb:results}
\input{includes/performance.tex}
\end{table}

\subsection{Ranking quality}
  \begin{figure*}[h]
    \centering
    \captionsetup{width=.8\linewidth}
    \caption{Maximum test MRR vs number of hyperparameter search iterations. Sequence-based models in the top row; factorization-based models in the bottom row.}
    \label{fig:hyper}
    \includegraphics[width=0.8\textwidth]{includes/hyperparam_search.pdf}
  \end{figure*}
In sequence-based models, our M-LSTM model outperforms the baseline LSTM model on all datasets. The performance gains are particularly large on the Amazon (26\%) datset and Goodbooks (17\%) datasets, and smaller, but still meaningful, on the Movielens 10M dataset (10\%). This is consistent with our conjectures on the nature of the datasets. The Amazon dataset, spanning the entire catalog of Amazon products, benefits most from being able to model taste diversity. This is not surprising, given that a single Amazon user can interact with products from many disparate domains.

@@

\subsection{Number of taste components}
\input{includes/hyperparameters.tex}
We summarize the effect of increasing the number of taste mixture components in Table~\ref{tab:nummixtures}. The optimum number of components is dataset and algorithm dependent but, by and large, there is a dose-response relationship between the number of mixtures and recommendation quality: being able to represent more distinct user tastes yields better results.

>>>>

\begin{table}
  \caption{Dataset statistics. 95th/50th denotes the ratio of popularity of the 95th and 50th percentile of item popularity.}
  \label{tab:datasets}
\begin{tabularx}{\columnwidth}{lllrr}
\toprule
 Dataset       & Users   & Items   &   Density &   95th/50th \\
\midrule
 Movielens & 69,879  & 10,678  &    0.0134 &      7.42 \\
 Amazon        & 100,085 & 113,997 &    0.0003 &      5.67 \\
 Goodbooks     & 53,425  & 10,001  &    0.0112 &      1.41 \\
\bottomrule
\end{tabularx}
\end{table}

@@

\section{Results}
Our main results are summarized in Table \ref{tb:results}. In both sequence-based and factorization tasks, variants of our model achieve consistently higher performance than baseline models, with improvements ranging from 10\% to 69\%. The results are robust to optimizing the hyperparameters of our baselines, giving us ample confidence that the results can be replicated in production recommender systems.
\begin{table}
\caption{Mean reciprocal rank across all users/sequences in the test set. Note that due to differences in experimental protocol, results between sequence-based and factorization models are not directly comparable.}
\label{tb:results}
\begin{subtable}{\columnwidth}
\caption{Sequence models}
\begin{tabularx}{\columnwidth}{lrrr}
\toprule
 Model        &   Movielens &   Amazon &   Goodbooks \\
\midrule
 LSTM         &          0.0908 &   0.1502 &      0.1158 \\
 Mixture-LSTM &          \textbf{0.1001} &   \textbf{0.1889} &      \textbf{0.1358} \\
\bottomrule
\end{tabularx}
\end{subtable}
\hspace{\fill}
\begin{subtable}{\columnwidth}
\caption{Factorization models}
\begin{tabularx}{\columnwidth}{lrrr}
\toprule
 Model              &   Movielens &   Amazon &   Goodbooks \\
\midrule
 Factorization           &          0.1053 &   0.1001 &      0.0738 \\
 Projection Mixture &          \textbf{0.1097} &   0.0698 &      0.0712 \\
 Embedding Mixture  &          0.1026 &   \textbf{0.1696} &      \textbf{0.0853} \\
\bottomrule
\end{tabularx}
\end{subtable}

\end{table}

\subsection{Ranking quality}
  \begin{figure*}[h]
    \centering
    \captionsetup{width=.8\linewidth}
    \caption{Maximum test MRR vs number of hyperparameter search iterations. Sequence-based models in the top row; factorization-based models in the bottom row.}
    \label{fig:hyper}
    \includegraphics[width=0.8\textwidth]{hyperparam_search.pdf}
  \end{figure*}
In sequence-based models, our M-LSTM model outperforms the baseline LSTM model on all datasets. The performance gains are particularly large on the Amazon (26\%) datset and Goodbooks (17\%) datasets, and smaller, but still meaningful, on the Movielens 10M dataset (10\%). This is consistent with our conjectures on the nature of the datasets. The Amazon dataset, spanning the entire catalog of Amazon products, benefits most from being able to model taste diversity. This is not surprising, given that a single Amazon user can interact with products from many disparate domains.

@@

\subsection{Number of taste components}
\begin{table}
\caption{Effect of number of mixture components}
\label{tab:nummixtures}
\begin{subtable}{\columnwidth}
\caption{Sequence models}
\begin{tabularx}{\columnwidth}{rrrr}
\toprule
   Components &   Movielens 10M &   Amazon &   Goodbooks \\
\midrule
            2 &          0.0882 &   0.1538 &      0.1262 \\
            4 &          0.0997 &   0.1689 &      0.1304 \\
            6 &          0.0922 &   \textbf{0.1889} &      \textbf{0.1358} \\
            8 &          \textbf{0.1001} &   0.1865 &      0.1327 \\
\bottomrule
\end{tabularx}
\end{subtable}
\hspace{\fill}
\begin{subtable}{\columnwidth}
\caption{Factorization models}
\begin{tabularx}{\columnwidth}{rrrr}
\toprule
   Components &   Movielens 10M &   Amazon &   Goodbooks \\
\midrule
            2 &          0.0931 &   0.1343 &      0.0797 \\
            4 &          \textbf{0.1026} &   0.1583 &      0.0840 \\
            6 &          0.0864 &   0.1612 &      0.0811 \\
            8 &          0.0879 &   \textbf{0.1696} &      \textbf{0.0853} \\
\bottomrule
\end{tabularx}
\end{subtable}\end{table}
We summarize the effect of increasing the number of taste mixture components in Table~\ref{tab:nummixtures}. The optimum number of components is dataset and algorithm dependent but, by and large, there is a dose-response relationship between the number of mixtures and recommendation quality: being able to represent more distinct user tastes yields better results.
